D.I.A.G.R.A.M.: Development of Image Analysis for Graph Recognition And Modeling

### 2025-05-08 doc/article/article.tex
Draft update

\section{Graph Diagram Recognition}

\subsection{Classifier Module}
\subsubsection{Preprocessing}
TODO
\subsubsection{Model}
TODO

\subsection{Extractor Module}
\subsubsection{Preprocessing}
TODO
\subsubsection{Bounding Box Detection}
TODO
\subsubsection{Content Recognition}

\subsection{Transducer Module}
\subsection{Compiler Module}

>>>>

% 
% Graph Diagram
%
\section{Graph Diagram Recognition}

\subsection{Internal Representation}
In order to pass data through the components of the pipeline in a unified way, it was necessary to decide upon an internal representation of the graph diagram. It consists of \textbf{elements} and \textbf{relations}. The latter are linked to at least one element. Every element has a \textit{category} and possibly some \textit{inner} and \textit{outer text}. Every relation has a \textit{category} too, either a \textit{source} or \textit{target element} or both, and possibly \textit{text} related to it.

This internal representation is generated from the image provided by the user by the Extractor module, while it is parsed and converted to a markup language of choice by the Transducer.

\subsection{Classifier Module}
\subsubsection{Preprocessing}
TODO \\

\subsubsection{Model}
TODO \\

\subsection{Extractor Module}
\subsubsection{Preprocessing}
TODO\\
\subsubsection{Bounding Box Detection}
TODO\\
\subsubsection{Content Recognition}
After having determined all the bounding boxes of the diagram image, in order to correctly create the diagram internal representation, there's the need to link together text, elements and relations.\\

\subsubsection{Text Digitization}
Since the task of converting an image of text into the relative string has already been solved in various ways, we settled upon the Microsoft \textit{Trocr Handwritten} Transformer model; more precisely, the \textit{small} version of it. [TODO: Fare esperimenti e provarne altri]. Every bounding box containing text that has been found is later passed to the model, and the resulting string generated by the model is linked to it.\\

\subsubsection{Text Association}
It can be assumed that every piece of text is linked to one and only one element or relation. So, every bounding box containing text is linked to the nearest element - it being either a relation or an element. The associations that require too long a distance are eliminated further in the algorithm, and the relevant text is discarded.\\

\subsubsection{Element-Text Association}
After that a text box has been assigned to an element, there's the need to understand if the text is either inside or outside the element and whether the text is actually referring to the given element or not. In order to do this, the overlap between the two bounding boxes is computed; if the overlap is over a certain threshold, heuristically decided, the association is kept. Otherwise, the distance between the two bounding boxes is computed - as the distance between their central point. If the distance is over a certain threshold, the association is kept; otherwise, it's discarded.\\

\subsubsection{Relation-Text Association}
After a text box is assigned to a relation, some computation is needed to fully understand how the text is associated to the relation itself. [TODO: Come trova source e target]. The relation bounding box is then split into three different sections. The overlap between the text bounding box and each of the three sections boxes is computed - then, the section with the highest overlap value is then assigned the text under consideration.
